---
title: "Project Proposal"
author: "MadagaskaR"
output: 
  html_document:
    highlight: tango
    theme: united
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    toc_depth: 3
date: "`r format(Sys.time(), '%d %B %Y')`"
---

## Group Members

- Barış SİVAS
- Neriman GÜRSOY
- Ozan Barış BAYKAN
- Sena KALEMCİ
- Tuğba ÜNAL



```{r message=FALSE, warning=FALSE, paged.print=TRUE}

library(readxl)
library(tidyverse)

raw_df <- read.csv("C:/Users/NERIMAN.GURSOY/Desktop/R data/earthquake.csv")

glimpse(raw_df)

```


## Overview of Dataset

The dataset covers up all the recorded earthquakes in Turkey between 1910 and 2017. There are 24007 rows and 17 variables. 

- id: id of earthquake
- date: date of earthquake
- time: time of earthquake
- lat: latitude of earthquake
- long: longitude of earthquake
- country: country of earthquake
- city: city of earthquake
- area: area of earthquake
- direction: direction of earthquake
- dist: distance of direction in km
- depth: depth of earthquake
- xm: xm of earthquake
- md: md of earthquake
- richter: Intensity of earthquake (Richter)
- mw: mw of earthquake
- ms: ms of earthquake
- mb: mb of earthquake

Dataset can be shown via [link](https://www.kaggle.com/caganseval/earthquake)

```{r }
summary(raw_df)
```

## Plan
